R Notebook

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
experiment with depth mesh

d <- temp_dat
response = "temperature_C_avg"
n_knots = 50
family = "gaussian"
plot = TRUE
fit_model = TRUE
extend = list(n = 8, offset = -0.1)
RangeGuess = 100 
offset = c(RangeGuess/5, RangeGuess)
max_edge = c(RangeGuess/5,RangeGuess)
cutoff = (RangeGuess/5)/5
prior_range = c(RangeGuess, 0.01) # P(range < 0.05) = 0.01
prior_sigma = c(0.5, 0.5)
kmeans = FALSE    
verbose = FALSE
debug = FALSE
trials = 2
covariates = c("depth_scaled", "depth_scaled2")
time = "year"
re_model_type = "rw1"
formula = full_model

fit_inla_sensor <- function(d, response = "temperature_C_avg",
                     n_knots = 50,
                     family = "binomial",
                     plot = FALSE,
                     fit_model = TRUE,
                     re_model_type = "rw1",
                     extend = list(n = 8, offset = -0.1),
                     kmeans = FALSE,                   
                     RangeGuess = 100, 
                     offset = c(RangeGuess/5, RangeGuess),                 
                     max_edge = c(RangeGuess/5,RangeGuess),
                     cutoff = (RangeGuess/5)/5,
                     prior_range = c(RangeGuess, 0.01), # P(range < 0.05) = 0.01
                     prior_sigma = c(0.5, 0.5),
                     time = "year",
                     covariates = NULL,
                     verbose = FALSE,
                     debug = FALSE,
                     trials = 10,
                     formula = full_model) {

@@

  ## Generate time mesh
  t_knots = seq(1, k, length = k)
  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")

@@

  # use kmeans() to calculate centers:
  if (kmeans) {
    km <- stats::kmeans(x = coords, centers = n_knots)
    mesh <- INLA::inla.mesh.create(km$centers,
      extend = extend
    )
  } else {
    bnd <- INLA::inla.nonconvex.hull(coords)

@@

@@

  ## build the space time indices
  # index = inla.spde.make.index("space", n.spde = spde$n.spde, n.group = mesh1$m)
  # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
  iset <- INLA::inla.spde.make.index(
    name = "spatial.field",
    n.spde = spde$n.spde, 
    n.group = mesh1$m
  )

  # isolate list of covariates:
  X.1              <- dat[, -c(1:4), drop = FALSE] # remove "y", "xcoo", "ycoo", "real_time"
  covar_names      <- colnames(X.1) # list covariates
  XX.list          <- as.list(X.1) # turn data.frame to a list
  effect.list      <- list()
  effect.list[[1]] <- c(iset)

 for (Z in seq_len(ncol(X.1)))
    effect.list[[Z + 1]] <- XX.list[[Z]]

  names(effect.list) <- c("1", covar_names)

 # projection matrix to map to observed points
 A <- INLA::inla.spde.make.A(
    mesh = mesh2d,
    loc = cbind(dat$xcoo, dat$ycoo),
    group = time_k, group.mesh = mesh1
  )
  A.list <- list()
  A.list[[1]] <- A

  for (Z in seq_len(ncol(X.1)))
    A.list[[Z + 1]] <- 1

@@

  full_model <- as.formula(paste(
    'y ~ -1 + ',
    paste(covar_names, collapse = '+'),
    '+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', re_model_type,'"))'))

  if (fit_model) {
    try_fitting <- TRUE
    trial_i <- 1
    while (try_fitting && trial_i <= trials) {
      model <- try(INLA::inla(formula,
                   family = family,
                   data = INLA::inla.stack.data(sdat),
                   control.predictor = list(compute = TRUE,
                                            A = INLA::inla.stack.A(sdat)),
                   control.fixed     = list(mean = 0, 
                                            prec = 1 / (5^2),
                                            mean.intercept = 0, 
                                            prec.intercept = 1 / (20^2)),
                   control.compute = list(config = TRUE),
                   verbose = verbose,
                   debug = debug,
                   keep = FALSE
                 )
      )
      trial_i <- trial_i + 1
      if (class(model) != "try-error") {
        try_fitting <- FALSE
      } else {
        # if (trial_i > 5) browser()
        message("INLA error. Trying again.")
      }
    }
    if (class(model) == "try-error") {
      model <- NA
    }
  } else {
    model <- NA
  }
  list(
    model = model, mesh = mesh, spde = spde, data = dat,
    formula = formula, iset = iset
  )
}
```

@@

median(pg_all$sp_field)
make_plot(pg_all, "sp_field") + facet_wrap(~ time) +
  scale_fill_gradient2(
    midpoint = median(pg_all$sp_field),
    high = scales::muted("blue"), 
    mid = "white",
    low = scales::muted("red"))

>>>>

d <- temp_dat
response = "temperature_C_avg"
n_knots = 50
family = "gaussian"
plot = TRUE
fit_model = TRUE
extend = list(n = 8, offset = -0.1)
RangeGuess = 100 
offset = c(RangeGuess/5, RangeGuess)
max_edge = c(RangeGuess/5,RangeGuess)
cutoff = (RangeGuess/5)/5
prior_range = c(RangeGuess, 0.01) # P(range < 0.05) = 0.01
prior_sigma = c(0.5, 0.5)
kmeans = FALSE    
verbose = FALSE
debug = FALSE
trials = 2
covariates = c("depth_scaled", "depth_scaled2")
time = "year"
sp_model_type = "ar1"
formula = "full_model"

fit_inla_sensor <- function(d, response = "temperature_C_avg",
                     n_knots = 50,
                     family = "binomial",
                     plot = FALSE,
                     fit_model = TRUE,
                     sp_model_type = "ar1",
                     extend = list(n = 8, offset = -0.1),
                     kmeans = FALSE,                   
                     RangeGuess = 100, 
                     offset = c(RangeGuess/5, RangeGuess),                 
                     max_edge = c(RangeGuess/5,RangeGuess),
                     cutoff = (RangeGuess/5)/5,
                     prior_range = c(RangeGuess, 0.01), # P(range < 0.05) = 0.01
                     prior_sigma = c(0.5, 0.5),
                     time = "year",
                     covariates = NULL,
                     verbose = FALSE,
                     debug = FALSE,
                     trials = 10,
                     formula = full_model) {

@@

@@

  # use kmeans() to calculate centers:
  if (kmeans) {
    km <- stats::kmeans(x = coords, centers = n_knots)
    mesh2d <- INLA::inla.mesh.create(km$centers,
      extend = extend
    )
  } else {
    bnd <- INLA::inla.nonconvex.hull(coords)

@@

  ## build the space time indices
  # index = inla.spde.make.index("space", n.spde = spde$n.spde, n.group = mesh1$m)
  # construct index for ar1 model 
  iset <- INLA::inla.spde.make.index(
    name = "spatial.field",
    n.spde = spde$n.spde, 
    n.group = k
  )

  ## Generate depth mesh
  d_knots = seq(min(d$depth_scaled), max(d$depth_scaled), length.out = 30)
  #mesh0 = inla.mesh.1d(loc = d_knots, degree = 0, boundary = c('neumann', 'free'))
  mesh1 = inla.mesh.1d(loc = d_knots, degree = 1, boundary = c('neumann', 'free'))
  #mesh2 = inla.mesh.1d(loc = d_knots, degree = 2, boundary = c('neumann', 'free'))
    # graphics::plot(mesh0$loc,mesh0$mid)
    # graphics::points(mesh1$loc,mesh1$mid)
    # graphics::points(mesh2$loc,mesh2$mid, col = "red")

  idepth <- INLA::inla.spde.make.index("depth.field", n.spde = mesh1$m, n.group = k)

@@

  # isolate list of covariates:
  X.1              <- dat[, -c(1:4), drop = FALSE] # remove "y", "xcoo", "ycoo", "real_time"
  covar_names      <- colnames(X.1) # list covariates
  XX.list          <- as.list(X.1) # turn data.frame to a list
  effect.list      <- list()
  effect.list[[1]] <- c(iset)
  effect.list[[2]] <- c(idepth)

 for (Z in seq_len(ncol(X.1)))
    effect.list[[Z + 2]] <- XX.list[[Z]]

  names(effect.list) <- c("1", "d_knots", covar_names)

 # projection matrix to map to observed points
 A <- INLA::inla.spde.make.A(
    mesh = mesh2d,
    loc = cbind(dat$xcoo, dat$ycoo),
    group = time_k
  )
 A2 <- INLA::inla.spde.make.A(
    mesh = mesh1,
    loc = dat$depth_scaled,
    group = time_k
  )
  A.list <- list()
  A.list[[1]] <- A
  A.list[[2]] <- A2

  for (Z in seq_len(ncol(X.1)))
    A.list[[Z + 2]] <- 1

@@

   full_model <- as.formula(paste(
    'y ~ -1 + ',
    paste(covar_names, collapse = '+'),
    '+f(depth_scaled, time_k, model = "rw1", scale.model = TRUE) + f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', sp_model_type,'"))'))

  if (fit_model) {
    try_fitting <- TRUE
    trial_i <- 1
    while (try_fitting && trial_i <= trials) {
      model <- try(INLA::inla(formula,
                   family = family,
                   data = INLA::inla.stack.data(sdat),
                   control.predictor = list(compute = TRUE,
                                            A = INLA::inla.stack.A(sdat)),
                   control.fixed     = list(mean = 0, 
                                            prec = 1 / (5^2),
                                            mean.intercept = 0, 
                                            prec.intercept = 1 / (20^2)),
                   control.compute = list(config = TRUE),
                   verbose = verbose,
                   debug = debug,
                   keep = FALSE
                 )
      )
      trial_i <- trial_i + 1
      if (class(model) != "try-error") {
        try_fitting <- FALSE
      } else {
        # if (trial_i > 5) browser()
        message("INLA error. Trying again.")
      }
    }
    if (class(model) == "try-error") {
      model <- NA
    }
  } else {
    model <- NA
  }
#  obj <- 
    list(model = model, mesh = mesh2d, spde = spde, data = dat,
    formula = formula, iset = iset
  )
}
```

@@

median(pg_all$sp_field)
make_plot(pg_all, "sp_field") + facet_wrap(~ time) +
  scale_fill_gradient2(
    midpoint = median(pg_all$sp_field),
    high = scales::muted("red"), 
    mid = "white",
    low = scales::muted("blue"))
